Formal Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2022-11-04 nakayama/paper.tex
Finished draft of math background, moving to informal NAK description

%\author{}
\author{

	Authors

	\vspace{.2cm}

	\\

	Institute for Information Sciences \\
	The University of Kansas \\
	Lawrence, KS 66045}
%\email{}

@@

\begin{abstract}
\emily[inline]{	Abstract here }
\end{abstract}

\noindent \textbf{Keywords:}
Formalization of Mathematics,
Formal Proof,
Commutative Algebra,
Commutative Ring,
Ideal,
Module over a Ring.

@@

Our goal is to newly formalize theoretical, rather than computational, commutative algebra in Coq.  We formally prove \emph{Nakayama's lemma} \cite{nakayama-1951, azumaya}, a fundamental theorem in the field. 
In the process of doing so,  we formalize certain algebraic structures  essential to higher-level algebra, including \emph{local rings} and \emph{modules over commutative rings}.

@@

\subsection{The Fundamental Algebraic Objects}

\emily[inline]{Something about building up a lot}

@@

The set of integers also forms a \emph{ring} because it is closed under multiplication, which again satisfies associativity, and the distributive law governing the compatibility of addition and multiplication holds. 
Even more, the integers form a 
\emph{commutative ring} since $n \cdot m = m \cdot n$ for all integers $n$ and $m$.  Moreover,  the integer $1$ is the ring's \emph{multiplicative identity}, since for any  $n \in \mathbb{Z}$ one has $n \cdot 1 = 1 \cdot n = n$, $\mathbb{Z}$ is, furthermore, a \emph{commutative ring with identity}.

In general, a \emph{ring} is a set $R$ with two binary operations, which we call \emph{addition} and \emph{mutiplication}, typically denoted $\cdot$ and $+$, respectively.  As motivated by the properties of the ring of integers, addition, $R$ must be an abelian group, multiplication must be associative, and the distributive law must hold, i.e., for all $x, y, z \in R$, $(x+y)\cdot z = x \cdot z + y \cdot z$ and $x \cdot (y+z) = x \cdot y + x \cdot z$.

\emily[inline]{Quickly mention familiar examples}

We will use $0$ to refer to an additive identity, and $1$ will represent a multiplicative identity.

Rather than call upon the \emph{Mathematical Components Library} 
 for Coq \cite{mathcomp}

@@

One can verify that given any integer $n$, the set $n\mathbb{Z}$ of its multiples forms an ideal of $\mathbb{Z}$.  For instance, $2 \mathbb{Z}$ consists of all even numbers, and is an abelian group under addition: the sum of two even numbers is even, the additive identity $0$ is eve, and the negative of an even number is even.
Moreover, the absorption property holds since the product of any integer and an even number is again even. 
In fact, every ideal of the ring of integers has this form $n\mathbb{Z}$ for some integer $n$, though ideals in general commutative rings can have more complicated properties.

@@

A \emph{maximal ideal} of a commutative ring is a proper ideal that is maximal with respect to inclusion, i.e., no other proper ideal strictly contains it. 
Going back to the ring of integers, $6 \mathbb{Z} \subsetneq 2 \mathbb{Z}$ since every multiple of $6$ is even, 
so $6 \mathbb{Z}$ is not a maximal ideal of $\mathbb{Z}$. 
However, no proper ideal $I$ contains $2 \mathbb{Z}$:  If $2 \mathbb{Z}\subsetneq I \subsetneq \mathbb{Z}$, then $I$  would necessarily contain an odd number $n$.  Writing $n=2k+1$ for some integer $k$, we notice that since $-2k = 2 \cdot -k$  is in $2\mathbb{Z}$, it is also an element of the larger set $I$, and since $I$ is an abelian group under addition, $(2k+1) + $

In fact, $3 \mathbb{Z}$

@@

>>>>

%\author{}
\author{ 
	Authors  
%	
	\vspace{.2cm}		
	\\ 
%	
	Institute for Information Sciences \\
	The University of Kansas \\
	Lawrence, KS 66045
	}
%\email{}

@@

\begin{abstract}
\emily[inline]{Abstract here }
\end{abstract}

\noindent \textbf{Keywords:}
Formalization of Mathematics,
Formal Proof,
Commutative Algebra,
Commutative Ring,
Local Ring,
Ideal,
Module over a Ring,
Finitely Generated Module.

@@

Our goal is to newly formalize theoretical, rather than computational, commutative algebra in Coq. 
We formally prove \emph{Nakayama's lemma} \cite{nakayama-1951, azumaya}, an essential result in the field. 
In doing so, we formalize algebraic structures that are fundamental to higher-level algebra, such as 
\emph{local rings} and \emph{(finitely generated) modules over commutative rings}. 
Rather than build upon some of the basic objects from abstract algebra, such as groups and rings, that have been formalized in Coq, e.g., in the  
\emph{Mathematical Components Library} 
\cite{mathcomp}, we start from scratch.

@@

\subsection{The Fundamental Algebraic Structures}

Here, we touch upon the mathematical structures from commutative algebra that are formalized.

@@

The set of integers also forms a \emph{ring} due to its properties of multiplication.  
It is closed under this binary operation, which satisfies associativity.  
The integer $1$ is a   \emph{multiplicative identity} since for any  $n \in \mathbb{Z}$ one has $n \cdot 1 = 1 \cdot n = n$.  Finally, the the distributive law governing the compatibility of addition and multiplication holds. 
Even more, the integers form a 
\emph{commutative ring} since $n \cdot m = m \cdot n$ for all integers $n$ and $m$.

In general, a commutative ring is a set $R$ with two binary operations, which we call \emph{addition} and \emph{mutiplication}, typically denoted $\cdot$ and $+$, respectively.  As motivated by the properties of the ring of integers, addition, $R$ must be an abelian group, multiplication must be associative, $R$ must have a multiplicative identity, and the distributive law must hold, i.e., for all $r, s, t \in R$, $(r+s)\cdot t = r \cdot t + s \cdot t$ and $r \cdot (s+t) = r \cdot s + r \cdot t$.

Other familiar examples of commutative rings include the integers modulo a fixed integer $n>0$, fields--commutative rings in which every nonzero element has a multiplicative inverse--such as the rings of rational numbers, real numbers, and complex numbers, and rings of polynomials in a variable $x$ with integer coefficients, or with coefficients in a field.

@@

One can verify that given any integer $n$, the set $n\mathbb{Z}$ of its multiples forms an ideal of $\mathbb{Z}$.  For instance, $2 \mathbb{Z}$ consists of all even numbers, and is an abelian group under addition: the sum of two even numbers is even, the additive identity $0$ is even, and the negative of an even number is even.
Moreover, the absorption property holds since the product of any integer and an even number is again even. 
In fact, every ideal of the ring of integers has this form $n\mathbb{Z}$ for some integer $n$, though ideals in general commutative rings can have more complicated properties.

@@

A \emph{maximal ideal} of a commutative ring is a proper ideal that is maximal with respect to inclusion, i.e., no other proper ideal strictly contains it. 
Returning to our example of the ring of integers, $6 \mathbb{Z} \subsetneq 2 \mathbb{Z}$ since every multiple of $6$ is even, 
so $6 \mathbb{Z}$ is not a maximal ideal of $\mathbb{Z}$. 
However, no proper ideal $I$ contains $2 \mathbb{Z}$:  If $2 \mathbb{Z}\subsetneq I \subsetneq \mathbb{Z}$, then $I$  would necessarily contain an odd number $n$.  Writing $n=2k+1$ for some integer $k$, we notice that since $-2k$ is in $2\mathbb{Z}$, it is also an element of the larger set $I$, and since $I$ is an abelian group under addition, $(2k+1) + (-2k) = 1$ is also in the ideal $I$. 
However, in this case, every integer $n = n \cdot 1$ is in $I$ by absorption, so $I = \mathbb{Z}$ is not a proper ideal, a contradiction.

In fact, $3 \mathbb{Z}$ is the only other maximal ideal of $\mathbb{Z}$ containing $6 \mathbb{Z}$, and in general, the prime ideals in the ring of integers besides the zero ideal are those of the form $p \mathbb{Z}$, where $p$ a prime number.

A commutative ring is \emph{local} if it has exactly one maximal ideal.  
Every field is local since the only proper ideal of a field is the zero ideal. 
By our observations above, the ring of integers is not local. 
However, the set of all rational numbers that can be written with an odd denominator does form a subring of all rational numbers, and its unique maximal ideal consists of the elements with even numerator; in fact, this ring is the so-called \emph{localization} of $\mathbb{Z}$ at the maximal ideal $2 \mathbb{Z}$.  
The ring of polynomials over a field in a variable $x$ is not local, but the ring of all formal power series is local; its maximal ideal consists of the power series with no constant term. 
The ring of integers modulo $n>1$ is local if and only if $n$ is a power of a prime number $p$,  in which case the unique maximal ideal consists of all multiples of $p$.

%Hence the ring of integers is not local, but 
%the ring of integers modulo an integer $n>1$,  often denoted $\mathbb{Z}/n\mathbb{Z}$ or $\mathbb{Z}_n$, are local

\paragraph{A module over a commutative ring.}  Let  $R$ be a commutative ring. 
A \emph{module over $R$}, or \emph{$R$-module}, is an abelian group $M$ under a binary operation $+$, and a {scalar multiplication} $R \times M \to M$ denoted $\cdot$, satisfying the following compatibility properties for all $r, s \in R$ and $u, v \in M$.

\begin{enumerate}[leftmargin=2cm,topsep=0cm,itemsep=0cm]
\item  $r \cdot (u + v)  = r \cdot u + r \cdot v$
\item $(r+s) \cdot u = r \cdot u + s \cdot u$
\item $(rs) \cdot u = r \cdot (s \cdot u)$
\item $1 \cdot u = u$ 
\end{enumerate}

From this definition, one can see that a module over a field $F$ is precisely an $F$-vector space, so the notion of a module over an arbitrary commutative ring extends that of a vector space over a field. 
Finitely generated vector spaces form the foundation for matrix algebra, and the extension of this notion to module theory is needed to state Nakayama's Lemma.  We call an $R$-module $M$ \emph{finitely generated} if there exist a fixed finite number of elements $u_1, \ldots, u_n$ of $M$ with the following property:  Given any  $w \in M$, there exist $r_1, \ldots, r_n \in R$ for which  
\[
w = r_1 u_1 + r_1 u_2 + \cdots + r_n u_n.
\]
When $R=F$ is a field and $M=V$ is a finite-dimensional vector space over $F$, one can choose $u_1, \ldots, u_n$ to be a basis for $V$, i.e., $n = \dim V$.  In this case, the choice of scalar coefficients in the expression above for $w \in V$ is unique.  When $R$ is not a field, modules can have more subtle properties, and such expression is typically not unique.

@@

In order to state Nakayama's Lemma, we clarify some notation:  If $I$ is an ideal of a commutative ring $R$ and 
 $M$ is an $R$-module, then $IM$ is the set of elements of the form 
	$a_1 u_1 + a_2 u_2 + \cdots + a_k u_k$, where $a_1, \ldots, a_k \in I$ and $u_1, \ldots, u_k \in M$.  
Notice that by the absorption property of ideals, $IM$ is an $R$-module contained in $M$.
